---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
if(!file.exists("activity.csv")) {
    unzip("activity.zip", c("activity.csv"))
}
data <- read.csv("activity.csv", header = T, sep = ",", na.strings = "NA")
data <- transform(data, date = as.Date(date, "%Y-%m-%d"))
str(data)
```

## What is mean total number of steps taken per day?
1. Total number of steps taken per day
```{r}
stepsPerDay <- aggregate(steps ~ date, data, sum, na.rm = T)
stepsPerDay
```

2. Histogram of the total number of steps taken each day
```{r}
hist(stepsPerDay$steps, main = NULL, xlab = NULL, ylab = NULL)
```

3. Mean and median of the total number of steps taken per day
```{r}
summary(stepsPerDay$steps)
```

## What is the average daily activity pattern?
1. Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
stepsPerInterval <- aggregate(steps ~ interval, data, mean, na.rm = T)
plot(stepsPerInterval$interval, stepsPerInterval$steps, type = "l", xlab = "Interval", ylab = "Average steps")
```

2. on average across all the days in the dataset, the 08:35 5-minute interval contains the maximum number of steps
```{r}
head(stepsPerInterval[order(stepsPerInterval$steps, decreasing = T), ], 5)
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
