---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
if(!file.exists("activity.csv")) {
    unzip("activity.zip", c("activity.csv"))
}
data <- read.csv("activity.csv", header = T, sep = ",", na.strings = "NA")
data <- transform(data, date = as.Date(date, "%Y-%m-%d"))
str(data)
```

## What is mean total number of steps taken per day?
1. Total number of steps taken per day
```{r}
stepsPerDay <- aggregate(steps ~ date, data, sum, na.rm = T)
stepsPerDay
```

2. Histogram of the total number of steps taken each day
```{r}
hist(stepsPerDay$steps, main = NULL, xlab = NULL, ylab = NULL)
```

3. Mean and median of the total number of steps taken per day
```{r}
summary(stepsPerDay$steps)
```

## What is the average daily activity pattern?
1. Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
stepsPerInterval <- aggregate(steps ~ interval, data, mean, na.rm = T)
plot(stepsPerInterval$interval, stepsPerInterval$steps, type = "l", xlab = "Interval", ylab = "Average steps")
```

2. On average across all the days in the dataset, the 08:35 5-minute interval contains the maximum number of steps
```{r}
head(stepsPerInterval[order(stepsPerInterval$steps, decreasing = T), ], 5)
```

## Imputing missing values
1. Analysis of total number of missing values in the dataset
```{r}
missingValues <- nrow(data) - sum(complete.cases(data))
missingValues
missingValues / nrow(data)
sum(is.na(data$steps))
sum(is.na(data$date))
sum(is.na(data$interval))
```

2. I´m filling the missing values with the mean for that 5-minute interval and create a new dataset
```{r}
avgStepsPerInterval <- aggregate(steps ~ interval, data, mean, na.rm = T)
avgStepsPerInterval <- transform(avgStepsPerInterval, steps = round(steps, digits = 0))
r <- merge(data, avgStepsPerInterval, by="interval", suffixes=c(".data", ".avg"))
na.idx <- which(is.na(data$steps))
completeData <- data
completeData[na.idx,"steps"] <- r[na.idx,"steps.avg"]
nrow(completeData) - sum(complete.cases(completeData))
```

3. The initial datased had 13.11% of missing values. After imputing missing values I observed increases in the mean and median. The mean changed from 10770 to 10890 (1.11% increase) and the median changed from 10760 to 11020 (2,41% increase).
```{r}
newStepsPerDay <- aggregate(steps ~ date, completeData, sum)
par(mfrow = c(1, 2))
hist(newStepsPerDay$steps, main = "With imputed missing values", xlab = NULL, ylab = NULL)
hist(stepsPerDay$steps, main = "With missing values", xlab = NULL, ylab = NULL)
summary(newStepsPerDay$steps)
summary(stepsPerDay$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
1. Creating a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
Sys.setlocale("LC_TIME", "C")
completeData[, "type"] <- factor(ifelse(weekdays(data$date, abbreviate = T) %in% c("Sat", "Sun"), "weekend", "weekday"))
str(completeData)
```

2. Created a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
par(mfrow = c(2, 1))
sub <- subset(completeData, type == "weekend")
steps <- aggregate(steps ~ interval, sub, mean)
plot(steps$interval, steps$steps, type = "l", xlab = "Interval", ylab = "Average steps", main = "Weekend")
sub <- subset(completeData, type == "weekday")
steps <- aggregate(steps ~ interval, sub, mean)
plot(steps$interval, steps$steps, type = "l", xlab = "Interval", ylab = "Average steps", main = "Weekday")
```